---
title: "My quarto part 2"
author: "Ally Malilay"
format: gfm
prefer-html: true
knitr: 
  opts_chunk: 
    fig.path: "../Output/"
---
#question: why doesn't my table appear on Quarto document as it shows in the qmd? 
I might have a problem with knitr:
# Introduction
Today, we are learning about how to work with figures.

# Load the libraries
```{r}
#| message: false
#| warning: false

library(here)
library(tidyverse)
library(palmerpenguins)
library(kableExtra)

```

# Make a plot
```{r}
#| warning: false
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-penguin
#| fig-cap: "This is a figure showing the flipper length by body mass"
penguins %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point()
```

When you look at @fig-penguin, we can see there is a positive relationship between flipper length and body mass.

| Time  | Session | Topic    |
|:------|:-------:|---------:|
| _left_ | _center_| _right_ |
| 01:00  | 1       | Anatomy |
| 01:50  |         | *Break* |
| 02:00  | 2       | Tables  |
| 02:45  |         | *Break* |

```{r}
#| label: tbl-penguin
#| tbl-cap: "My awesome penguin table"
penguins %>%
  group_by(species) %>%
  summarise(flipperlengthmean = mean(flipper_length_mm, na.rm = TRUE)) %>%
  kbl() %>% #add kable table
kable_classic() %>% #add theme
row_spec(2, bold = TRUE, color = "black", background = "pink") %>% #edit appearance of row 2
kable_styling(full_width = FALSE)
```

@tbl-penguin contains information on the average flipper length of the three penguin species.