---
title: "Tidy Tuesday 4"
format: html
toc: true
author: "Ally Malilay"
date: "2025-12-6"
theme: united
mainfont: "Arial, sans-serif"
---

# Load libraries and data
Today we're looking at data that monitors the statistical performance of countries as develped by the World Bank. The [SPI Framework](https://www.worldbank.org/en/programs/statistical-performance-indicators/Framework) focuses on five key dimensions of a country’s statistical performance: (i) data use, (ii) data services, (iii) data products, (iv) data sources, and (v) data infrastructure. 
```{r}
#| message: false
#| warning: false
#Libraries
library(tidytuesdayR)
library(ggplot2)
library(tidyr)
library(dplyr)
library(plotly)
library(scales)
library(htmlwidgets)
library(here)

# Read in Tidy Tuesday data
spi <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-11-25/spi_indicators.csv')
```

# Tidy up the data!
We're going to take a look at the most recent data from 2023. We want to understand the relationship between the data products score and data infrastructure score.
```{r}
#| message: false
#| warning: false

spi_cleaned <- spi %>%
  filter(year == "2023") %>% #most recent year
  select(-year) %>% #remove year column from data
  drop_na("data_products_score","data_infrastructure_score") %>% #drop NAs for data we're using
  filter(income != "Not classified") %>% #filter for income not equal to not classified
   mutate(income = factor(income, #factor for legend
          levels = c("High income", "Upper middle income", "Lower middle income", "Low income")),
          pop_group = cut(population, #factor for population, pop range huge &plotly legends can only handle discrete
                          breaks = c(0, 1000000, 100000000, 500000000, 1428627663),
                          labels = c("Under 1M", "50M - 100M", "100M - 500M", "Over 500M")
          ))
          

```

# Our first time making a plot with bubble graphs
Let's see how the data products and infrastructure scores change based on population size and income. Is there a clear relationship we can identify here?
```{r}
#| message: false
#| warning: false
plot <- ggplot(spi_cleaned, 
       aes(x = data_products_score, y = data_infrastructure_score,
                size = pop_group, color = income,
           text = paste("Country:", country, #text for interactive plotly
                         "<br>Data Products Score:", data_products_score,
                        "<br>Data Infrastructure Score:", data_infrastructure_score,
                         "<br>Population:", pop_group))) +
  geom_point(alpha = 0.5) + #modify color transparency
  scale_size_manual(
             name = "Population Size",
              values = c(1,5, 10,18)) +    
  scale_color_discrete(name = "Income") + #legend title
  labs(x = "Data Products Score",
       y = "Data Infrastructure Score",
       title = "Global Data Infrastructure Score vs Data Products Score in 2023",
       subtitle = "By country, population, and income") +
  facet_wrap(~region) 

plot  
ggsave(here("Tidy_Tuesday","2025-11-25","Output","GlobalDataScores.png"))
```

# Make the plot interactive with a new package!
It would be great to see the specific countries we're taking a look at and their specific data scores. 
```{r}
#| message: false
#| warning: false
GlobalData <- ggplotly(plot,
                        tooltip = c("text")) #for the popup text

GlobalData
saveWidget(GlobalData, here("Tidy_Tuesday","2025-11-25", "Output", "GlobalDataPlotly.html"))

```

# What we learned today
This was the first time that I used a bubble plot and plotly. I converted my data to factors for population and income levels for better data visualization given the huge range in population size and categorical nature of income levels. For a bubble plot, the transparent color was necessary since countries with larger populations covered the data of countries with smaller populations. I have the two different plots shown here because plotly legends are very finicky - I had to convert population to factors, otherwise it did not appear in the legend. The plotly legend is messy and the legend can't divide population size and income the way a ggplot legend can.